Analyzing The National's Musical Trajectory from 2001 to 2023

### 2024-10-10 paper/paper.qmd
final paper updates

Write Introduction here

@@

# Plot 1: danceability over time
ggplot(data, aes(x = album_release_year, y = danceability)) +
  geom_point(alpha = 0.7, color = "black") +
  stat_summary(fun = mean, geom = "line", color = "green", size = 1.2) +
  labs(
    x = "Album Release Year",
    y = "Danceability"
  ) +
  theme_minimal()
```

@@

# Plot 2: energy over time
ggplot(data, aes(x = album_release_year, y = energy)) +
  geom_point(alpha = 0.7, color = "black") +
  stat_summary(fun = mean, geom = "line", color = "blue", size = 1.2) +
  labs(
    x = "Album Release Year",
    y = "Energy"
  ) +
  theme_minimal()
```

@@

# Plot 3: instrumentalness over time
ggplot(data, aes(x = album_release_year, y = instrumentalness)) +
  geom_point(alpha = 0.7, color = "black") +
  stat_summary(fun = mean, geom = "line", color = "red", size = 1.2) +
  labs(
    x = "Album Release Year",
    y = "Instrumentalness"
  ) +
  theme_minimal()
```

@@

>>>>

In recent years, the music analytics field has advanced significantly, providing new insights into how artists' sounds evolve over time. This study delves into the discography of The National, a prominent indie rock band known for their rich, emotive soundscapes and introspective lyricism, by examining key musical attributes such as danceability, energy, instrumentalness, and valence across their albums. By utilizing data drawn from Spotify's API, we aim to track how these characteristics have transformed throughout the band's career, illustrating their progression with comprehensive visualizations. Our analysis reveals notable trends and variations, particularly pointing out that energy and instrumentalness exhibit considerable changes, perhaps reflecting shifts in the band’s artistic direction, experimentation, or production choices. Conversely, danceability and valence maintain relatively stable trajectories, possibly indicating consistency in their core aesthetic approach. These insights not only illuminate The National's evolving sound but also contribute to the broader conversation in music analytics, enhancing our understanding of artistic development and genre dynamics within the indie rock landscape. Through this study, we offer a deeper appreciation of how technological tools can be leveraged to explore and comprehend music's intricate evolution over time.

@@

# Plot 1: danceability over time
ggplot(data, aes(x = album_release_year, y = danceability)) +
  geom_point(alpha = 0.7, color = "black") +
  stat_summary(fun = mean, geom = "line", color = "green", size = 1.2) +
  labs(
    x = "Album Release Year",
    y = "Danceability"
  ) +
  theme_minimal()
```
\FloatBarrier

@fig-dance displays the danceability of music by The National since the band’s establishment. The National showed consistently high danceability throughout their career despite a visible drop after their initial 2001 album and their 2017 album. This trend reveals that The National had a stable style after it established itself as a rock band.

@@

# Plot 2: energy over time
ggplot(data, aes(x = album_release_year, y = energy)) +
  geom_point(alpha = 0.7, color = "black") +
  stat_summary(fun = mean, geom = "line", color = "blue", size = 1.2) +
  labs(
    x = "Album Release Year",
    y = "Energy"
  ) +
  theme_minimal()
```
\FloatBarrier

Energy was more volatile compared to danceability, as displayed in @fig-energy. In the 2001 and 2019 releases, the average energy levels were around 0.5, while in most years, energy levels were greater than 0.6. This demonstrates The Nationa’s willingness to experiment, though it held on to its rock identity by producing music with high overall energy.

@@

# Plot 3: instrumentalness over time
ggplot(data, aes(x = album_release_year, y = instrumentalness)) +
  geom_point(alpha = 0.7, color = "black") +
  stat_summary(fun = mean, geom = "line", color = "red", size = 1.2) +
  labs(
    x = "Album Release Year",
    y = "Instrumentalness"
  ) +
  theme_minimal()
```
\FloatBarrier

Instrumentalness was the most volatile aspect of The National’s musical style over time. As shown in @fig-instrumental. We can identify that the band experimented with very high instumentalness in 2007 and 2017, while the majority of songs released in other years showed a moderate level of instrumentalness. Combined with other trends, we can infer that The National is more willing to attempt music with different compositions, such as songs with more instrumentals and fewer lyrics, but unwilling to shift its fundamental rock style.

@@

Lastly, as indicated in @fig-valence, The Nationals maintained a moderate level of valence throughout the years, consistent with the observation that the band strived to keep its fundamental identity.
